---
title: "sap_landscape_setup"
format: html
editor: visual
---

## Landscape Setup

This document features code to setup landscapes for simulations.

### Setup

First load the requisite libraries, as well as the script featuring the landscape percolation function.

```{r}
library(readr) # for writing rds
library(future)
library(furrr) # for parallel computing
source("parasite_sim_functions.R") ## loads required packages as well
```

### Percolation

This code block creates the parameter space (parasite density as proportion of cells filled, parasite spatial aggregation as cluster factor) and creates a landscape for each parameter combo. Parallel computing is used for speed, since the landscape percolation algorithm is a bit slow.

```{r}
## save proportions (parasite densities)
props <- rep((1:9)/10, 16)

## save clustering inputs
clusters <- rep(0:15, each = 9)

## setup multisession computing
plan(multisession)

## percolate landscapes
ls_list <- future_map2(
  .x = props,
  .y = clusters,
  ~fast_ls_perc(size = 48, prop = .x, cluster = .y),
  .options = furrr_options(seed = TRUE),
  .progress = TRUE)
```

Save as an rds file for easy access within R:

```{r}
timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")
write_rds(ls_list, str_c("sim_output/ls_list_", timestamp, ".rds"))
```
