---
title: "sap_hoover_index_analysis"
format: html
editor: visual
---

## Analyzing Simulation Results with the Hoover Index

This script calculates the Hoover index value for each simulation of a set. Hoover index varies from 0 to one, representing the proportion of parasites needed to be rearranged to generate an even distribution among hosts (0, means already even, \~1 means highly aggregated). See McVinish and LEster (2020) "Measuring aggregation in parasite populations"

### Setup

```{r}
## load code
library(tidyverse)

## load sim data
recov_sim_data <- read_rds("sim_output/recov_sim_data20230417_132954.rds")
```

### Calculate

```{r}

## create a hoover index function

## make dummy parasite burden data
burdens <- rpois(n = 96, lambda = 10)

## sketch out steps
mean_burden <- mean(burdens)
hoover <- sum(abs(burdens - mean_burden))/(2*length(burdens)*mean_burden)

## write generic function
hoover_index <- function(values){
  mean_val <- mean(values)
  hoover <- sum(abs(values - mean_val))/(2*length(values)*mean_val)
  return(hoover)
}


## test function for a single simulation

recov_sim_data <- mutate(recov_sim_data,
                         sim_id = str_c(prop, cluster, recov_prob, rep_id, sep = "_"))
single_sim <- filter(recov_sim_data, sim_id == "0.1_0_0.05_1")

test_hi <- hoover_index(single_sim$cur_burden)

## do for all sims
summary_data <- recov_sim_data %>%
  group_by(prop, cluster, moran, rep_id, recov_prob) %>%
  summarize(mean_cur = mean(cur_burden),
            mean_cum = mean(cum_burden),
            var_cur = var(cur_burden),
            var_cum = var(cum_burden),
            disp_cur = var_cur / mean_cur,
            disp_cum = var_cum / mean_cum,
            disp_diff = disp_cur - disp_cum,
            mean_diff = mean_cur - mean_cum,
            var_diff = var_cur - var_cum,
            hoover = hoover_index(cur_burden))


```

### Visualize

```{r}

##plot hoover index

fig1 <- ggplot(summary_data, aes(x = moran, y = hoover, color = factor(prop))) +
  geom_point(size = 4) +
  geom_smooth(se = FALSE, method = "lm", linewidth = 3) +
  facet_wrap(vars(recov_prob)) +
  scale_color_viridis_d(name = "Parasite Density \n(Proportion Cells Occupied)") +
  labs(y = "Hoover Index of \n'Snapshot' Burden Distribution") +
  scale_x_continuous(name = "Parasite Clustering (Moran's I)",
                     breaks = c(0, 0.06, 0.12, 0.18),
                     sec.axis = sec_axis(~ . , name = "Recovery Rate", 
                                         breaks = NULL, labels = NULL)) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.title = element_text(size = 24),
        legend.text = element_text(size = 16),
        strip.text = element_text(size = 24),
        axis.title = element_text(size = 32),
        axis.text = element_text(size = 16),
        axis.title.x.top = element_text(size = 32, vjust = 1))


## save hoover index plot

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")

## save fig 1
ggsave(str_c("figures/sdp_ms_fig3_", timestamp, ".tiff"), plot = fig1,
       units = "in", width = 12, height = 8)
```

### Checking distributions

```{r}

recov_sim_data %>%
  filter(prop == 0.5, cluster == 15) %>%
  ggplot(aes(x = cur_burden)) +
  geom_histogram() +
  facet_wrap(vars(recov_prob))
```
