---
title: "sdp_individual_move_variation_nb"
output: html_document
---

### Individual Variation of Movement

First, load libraries

```{r}
library(tidyverse)
library(som.nn)
library(ape)
```

#### Revise Move Function

Change the potential moves from sampling separate x and y values to sampling pairs, thus the ratios can be modified

```{r}

move_host <- function(hid, hosts, coords, n_moves, inactivity = 1){
  ## grab host
  host_start <- filter(hosts, host_id %in% hid)
  
  ## grab landscape size for torus wrap
  max_x <- max(coords$x)
  max_y <- max(coords$y)
  
  ## set up moves to sample from
  moveset <- tibble(
    x = c(-1, -1, -1, 0, 0, 1, 1, 1, rep(0, inactivity)),
    y = c(-1, 0, 1, -1, 1, -1, 0, 1, rep(0, inactivity))
    )
  
  ## random moves
  moves <- slice_sample(moveset, n = n_moves, replace = TRUE)
  xmoves <- moves$x
  ymoves <- moves$y
  
  ## calculate x positions
  xpos <- cumsum(c(host_start$x, xmoves))
  
  ## calculate y positions
  ypos <- cumsum(c(host_start$y, ymoves))
  
  ## add positions to data
  host_moved <- tibble(host_id = hid,
                       xmove = c(NA, xmoves),
                       ymove = c(NA, ymoves),
                       pre_x = xpos,
                       pre_y = ypos,
                       time = 0:n_moves) %>%
    mutate(oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
           oob_y = (pre_y - 1) %/% max_y,
           x = pre_x - (max_x * oob_x), ## wrap around torus
           y = pre_y - (max_y * oob_y),
           id = str_c(x, y, sep = "_")) %>% ## generate ids
    left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
    mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) ## attach parasites
  
  ## return sim data
  return(host_moved)
}

```

Now test it

```{r}
inactivity = 1
moveset <- tibble(
    x = c(-1, -1, -1, 0, 0, 1, 1, 1, rep(0, inactivity)),
    y = c(-1, 0, 1, -1, 1, -1, 0, 1, rep(0, inactivity))
    )

size <- 12
coords <- tibble(x = rep(1:size, size), 
                   y = rep(1:size, each = size),
                   id = str_c(x, y, sep = "_"),
                   weight = rep(1, size^2), # Weight starts equal for all cells
                   value = rep(0, size^2)) # Cell type starts out as 0 for all

coords$value <- rbinom(n = size^2, size = 1, prob = 0.25)

n_hosts <- 24
host_origins <- slice_sample(coords, n = n_hosts, replace = TRUE)
  
hosts <- host_origins %>%
  select(x, y, id) %>%
  mutate(host_id = seq_len(nrow(host_origins)),
         time = 0)

test_host <- move_host(hid = 1, hosts = hosts, coords = coords, n_moves = 100, inactivity = 10)

test_inactivity <- map_df(1:16, 
                          function(x){
                            dat <- move_host(hid = 1, hosts = hosts, coords = coords,
                                             n_moves = 300, inactivity = x)
                            spaces <- length(levels(factor(dat$id)))
                            return(tibble(inactivity = x, spaces = spaces))
                          })

qplot(x = inactivity, y = spaces, data = test_inactivity)
```

Works great! Increasing "inactivity" make the host move less, visit fewer spaces

#### More Movement Optimization

However, the above function still seems inefficient as it does things one host at a time.

```{r}

size <- 12
coords <- tibble(x = rep(1:size, size), 
                   y = rep(1:size, each = size),
                   id = str_c(x, y, sep = "_"),
                   weight = rep(1, size^2), # Weight starts equal for all cells
                   value = rep(0, size^2)) # Cell type starts out as 0 for all

coords$value <- rbinom(n = size^2, size = 1, prob = 0.25)

n_hosts <- 24
host_origins <- slice_sample(coords, n = n_hosts, replace = TRUE)
n_moves <- 300
  
hosts <- host_origins %>%
  select(x, y, id) %>%
  mutate(host_id = seq_len(nrow(host_origins)),
         inactivity = sample(c(0, 0.25, 0.5, 0.75), size = n(), replace = TRUE))



## grab landscape size for torus wrap
max_x <- max(coords$x)
max_y <- max(coords$y)
  
## set up moves to sample from
moveset <- tibble(
    move_id = 1:9,
    xmove = rep(c(0, -1, 1), 3),
    ymove = rep(c(0, -1, 1), each = 3)
    )
  
c(1, rep(5, 0))

## random moves


##### THIS IS IT 
hosts_moving <- hosts[rep(1:nrow(hosts), each = n_moves + 1),] %>%
  mutate(time = rep(0:n_moves, n_hosts),
         active = inactivity * runif(n()) >= inactivity^2,
         move_id = sample(2:9, size = n(), replace = TRUE)^active) %>%
  left_join(moveset, by = "move_id") %>%
  mutate(xmove = if_else(time == 0, x, as.integer(xmove)),
         ymove = if_else(time == 0, y, as.integer(ymove))) %>%
  group_by(host_id) %>%
  mutate(pre_x = cumsum(xmove),
         pre_y = cumsum(ymove),
         oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
         oob_y = (pre_y - 1) %/% max_y,
         x = pre_x - (max_x * oob_x), ## wrap around torus
         y = pre_y - (max_y * oob_y),
         id = str_c(x, y, sep = "_")
         ) %>%
  left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
  mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) ## attach parasites
#####

hosts_moving %>%
  group_by(host_id) %>%
  summarise(inactivity = mean(inactivity),
            moves = mean(active))

length(rep(c(NA, inactivity * sample(1:9, size = n_moves, replace = TRUE) < 1), n_hosts))

(9^FALSE)

1/9 * (1:10)

moves <- sample(c(1:9, rep(5, inactivity - 1)), n = n_moves, replace = TRUE)
xmoves <- moves$x[moves]
ymoves <- moves$y[moves]
  
## calculate x positions
xpos <- cumsum(c(host_start$x, xmoves))
  
## calculate y positions
ypos <- cumsum(c(host_start$y, ymoves))
  
## add positions to data
host_moved <- tibble(host_id = hid,
                       xmove = c(NA, xmoves),
                       ymove = c(NA, ymoves),
                       pre_x = xpos,
                       pre_y = ypos,
                       time = 0:n_moves) %>%
    mutate(oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
           oob_y = (pre_y - 1) %/% max_y,
           x = pre_x - (max_x * oob_x), ## wrap around torus
           y = pre_y - (max_y * oob_y),
           id = str_c(x, y, sep = "_")) %>% ## generate ids
    left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
    mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) ## attach parasites
```

now let's put that in a function

```{r}
move_host_fast <- function(hosts, coords, n_moves){
  
  ## grab landscape size for torus wrap
  max_x <- max(coords$x)
  max_y <- max(coords$y)
  
  ## set up moves to sample from
  moveset <- tibble(
      move_id = 1:9,
      xmove = rep(c(0, -1, 1), 3),
      ymove = rep(c(0, -1, 1), each = 3)
      )
  
  hosts_moving <- hosts[rep(1:nrow(hosts), each = n_moves + 1),] %>%
    mutate(time = rep(0:n_moves, n_hosts),
           active = inactivity * runif(n()) >= inactivity^2,
           move_id = sample(2:9, size = n(), replace = TRUE)^active) %>%
    left_join(moveset, by = "move_id") %>%
    mutate(xmove = if_else(time == 0, x, as.integer(xmove)),
           ymove = if_else(time == 0, y, as.integer(ymove))) %>%
    group_by(host_id) %>%
    mutate(pre_x = cumsum(xmove),
           pre_y = cumsum(ymove),
           oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
           oob_y = (pre_y - 1) %/% max_y,
           x = pre_x - (max_x * oob_x), ## wrap around torus
           y = pre_y - (max_y * oob_y),
           id = str_c(x, y, sep = "_")
           ) %>%
    left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
    mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) ## attach parasites

    return(hosts_moving)
}


size <- 48
coords <- tibble(x = rep(1:size, size), 
                   y = rep(1:size, each = size),
                   id = str_c(x, y, sep = "_"),
                   weight = rep(1, size^2), # Weight starts equal for all cells
                   value = rep(0, size^2)) # Cell type starts out as 0 for all

coords$value <- rbinom(n = size^2, size = 1, prob = 0.25)

n_hosts <- 96 * 10 * 20
host_origins <- slice_sample(coords, n = n_hosts, replace = TRUE)
n_moves <- 300
  
hosts <- host_origins %>%
  select(x, y, id) %>%
  mutate(host_id = seq_len(nrow(host_origins)),
         inactivity = sample(c(0), size = n(), replace = TRUE))

Sys.time()
fast_moved <- move_host_fast(hosts = hosts, coords = coords, n_moves = 300)
Sys.time()

Sys.time()
moved <- map_df(1:n_hosts, 
                ~move_host(hid = .x, hosts = hosts, coords = coords, n_moves = 300),
                .progress = TRUE)
Sys.time()

```

Very fast!

Now minor edits to function for replication, no diagonal movement

```{r}
move_host_fast <- function(hosts, coords, n_moves, n_reps){
  
  ## grab landscape size for torus wrap
  max_x <- max(coords$x)
  max_y <- max(coords$y)
  
  ## set up moves to sample from
  moveset <- tibble(
      move_id = 1:5,
      xmove = c(0, 0, 0, -1, 1),
      ymove = c(0, -1, 1, 0, 0)
      )
  
  hosts_moving <- hosts[rep(1:nrow(hosts), each = n_moves + 1),] %>%
    mutate(time = rep(0:n_moves, n_hosts * n_reps),
           active = inactivity * runif(n()) >= inactivity^2,
           move_id = sample(2:5, size = n(), replace = TRUE)^active) %>%
    left_join(moveset, by = "move_id") %>%
    mutate(xmove = if_else(time == 0, x, as.integer(xmove)),
           ymove = if_else(time == 0, y, as.integer(ymove))) %>%
    group_by(rep_id, host_id) %>%
    mutate(pre_x = cumsum(xmove),
           pre_y = cumsum(ymove),
           oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
           oob_y = (pre_y - 1) %/% max_y,
           x = pre_x - (max_x * oob_x), ## wrap around torus
           y = pre_y - (max_y * oob_y),
           id = str_c(x, y, sep = "_")
           ) %>%
    left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
    mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) ## attach parasites

    return(hosts_moving)
}
```

#### Revise Sim Function

Now, allow the simulation function to pass inactivity values

```{r}
n_hosts <- 10
movetypes <- c(0.2, 0.1)
moveprops <- c(0.9, 0.1)
move_n <- n_hosts * moveprops

n_reps = 2
## Set random starting positions for each host
  host_origins <- slice_sample(coords, n = n_hosts * n_reps, replace = TRUE)
  
  ## Create host data frame
  hosts <- host_origins %>%
    select(x, y, id) %>%
    mutate(host_id = rep(seq_len(n_hosts), n_reps),
           time = 0,
           rep_id = rep(seq_len(n_reps), each = n_hosts),
           inactivity = rep(rep(movetypes, moveprops * n_hosts), n_reps))

  
move_host_fast(hosts = hosts, coords = coords, n_moves = 100, n_reps = 2)
  
parasite_sim <- function(landscape, n_hosts, n_moves, n_reps = 1, 
                         movetypes = c(0.2), moveprops = c(1)){
  
  ## Set random starting positions for each host
  host_origins <- slice_sample(landscape$coords, n = n_hosts * n_reps, replace = TRUE)
  
  ## Create host data frame
  hosts <- host_origins %>%
    select(x, y, id) %>%
    mutate(host_id = seq_len(nrow(host_origins)),
           time = 0,
           rep_id = rep(seq_len(n_reps), each = n_hosts),
           inactivity = rep(rep(movetypes, moveprops * n_hosts), n_reps))
  
  ## Simulate movement and parasite gain
  hosts_moved <- move_host_fast(hosts = hosts, coords = coords, 
                                n_moves = n_moves, n_reps = n_reps)
  
  ## Package and return host and landscape info
  out <- bind_cols(hosts_moved, landscape$ls_stats) %>%
    filter(time == n_moves)
  return(out)
}
```

That should work! Now make a test landscape

```{r}
fast_ls_perc <- function(size, prop, cluster){
  ## Calculate number of cells to fill based on size and prop
  potential <- floor((size^2) * prop)
  ## Create empty data representing the landscape
  coords <- tibble(x = rep(1:size, size), 
                   y = rep(1:size, each = size),
                   id = str_c(x, y, sep = "_"),
                   weight = rep(1, size^2), # Weight starts equal for all cells
                   value = rep(0, size^2)) # Cell type starts out as 0 for all
  
  ## Initialize the possible percolation locations
  possibles <- coords$id
  ## Percolate as many cells as need to be filled
  for(p in seq_len(potential)){
    
    ## Pick a random cell (weighted by the "weight" column) to set as landscape value "1"
    point <- sample(possibles, size = 1, prob = coords[coords$id %in% possibles,]$weight)
    coords[coords$id %in% point, "value"] <- 1
    
    ## Record current ones
    ones <- coords$value == 1
    
    ## Make an inverse distance matrix of all points
    inv_dists <- 1/as.matrix(dist(cbind(coords$x, coords$y)))
    ## Remove Inf values
    diag(inv_dists) <- 0
    
    ## Select only the columns of points with 1 as their value
    inv_dists_filt <- as.matrix(inv_dists[,ones])
    
    ## Assign the weights as each point's nearest distance to a 1 
    coords$weight <- map_dbl(1:nrow(inv_dists_filt), 
                             function(x){
                               (1 + max(inv_dists_filt[x,]))^cluster
                             })
    
    ## Refill the possible cells for percolation as only the remaining zeroes
    possibles <- coords %>%
      filter(value == 0) %>%
      #slice_max(weight, n = floor((length(coords$weight) * cluster))) %>%
      pull(id)
    
  }
  
  ## Create a distance matrix of each point assuming a torus shape
  ls_dists <- as.matrix(som.nn::dist.torus(cbind(coords$x, coords$y)))
  
  ## Invert the distance matrix
  ls_dists_inv <- 1/ls_dists
  diag(ls_dists_inv) <- 0
  
  ## Calculate Moran's I spatial autocorrelation from the inverse distance matrix
  moran_result <- Moran.I(coords$value, ls_dists_inv)
  
  ## Save the statistical output from the Moran test
  ls_stats <- tibble(size = size,
                     prop = prop,
                     cluster = cluster, 
                     moran = moran_result$observed, 
                     p = moran_result$p.value)
  
  ## Save a list of the coordinates data frame, a landscape matrix, and the Moran stats
  landscape <- list(coords = coords, 
                    matrix = matrix(coords$value, nrow = size, ncol = size),
                    ls_stats = ls_stats)
  
  ## Return that landscape list
  return(landscape)
}

test_ls <- fast_ls_perc(48, prop = 0.2, cluster = 5)

```

And then test the parasite sim

```{r}
n_moves = 300
n_hosts = 96
n_reps = 2
length(rep(0:n_moves, n_hosts * n_reps))

test_sims <- parasite_sim(landscape = test_ls, n_hosts = 96, n_moves = 300,
                          n_reps = 2, movetypes = c(0.2), moveprops = c(1))
```
