---
title: "sdp_parasite_sim_testing_nb"
output: html_document
---

### Parasite Simulation Testing

First, source the functions:

```{r}
library(tidyverse)
library(som.nn)
library(ape)
source("parasite_sim_functions.R")
```

#### Testing Variance of Outcomes on the Same Landscape

Now, how does the variance of characteristics of a population's parasite burden distribution change with number of timesteps?

First, generate a landscape:

```{r}
test_ls <- fast_ls_perc(size = 24, prop = 0.2, cluster = 15)
```

Then run a parasite simulation on it many times

```{r}

## just one
test_sim <- parasite_sim(landscape = test_ls, n_hosts = 48, n_moves = 1000)

## now many
test_sims <- replicate(10, 
                       parasite_sim(landscape = test_ls, n_hosts = 48, n_moves = 1000),
                       simplify = FALSE)

## make df with id
test_sims_df <- bind_rows(test_sims, .id = "sim_id")

## summarize by simulation and timepoint
test_summary <- test_sims_df %>%
  group_by(sim_id, time) %>%
  summarise(mean_burden = mean(parasite_burden),
            var_burden = var(parasite_burden),
            disp_burden = var_burden / mean_burden)

## plot summary stat by time
ggplot(data = test_summary) +
  geom_line(aes(x = time, y = disp_burden))

```

Looks like the variance among simulations doesn't decline with timesteps, but the dispersion does level off maybe around 200 timesteps?

But does that level off happen similarly for other clustering factors?

```{r}
test_ls <- fast_ls_perc(size = 24, prop = 0.2, cluster = 20)

## now many
test_sims <- replicate(10, 
                       parasite_sim(landscape = test_ls, n_hosts = 48, n_moves = 1000),
                       simplify = FALSE)

## make df with id
test_sims_df <- bind_rows(test_sims, .id = "sim_id")

## summarize by simulation and timepoint
test_summary <- test_sims_df %>%
  group_by(sim_id, time) %>%
  summarise(mean_burden = mean(parasite_burden),
            var_burden = var(parasite_burden),
            disp_burden = var_burden / mean_burden)

## plot summary stat by time
ggplot(data = test_summary) +
  geom_line(aes(x = time, y = disp_burden))
```

The level off point increases with clustering. Select an end point that for the max clustering used. Let's go with 300 for cluster = 20.

That means we want a trimmed movement sim function

#### Trimming Movement Simulation

Setting the movement function to only output the last timestep info

```{r}

#####
# Host Movement Function
##

#' Function inputs:
#' hid, id of the host
#' hosts, data frame of host info
#' coords, coords of the landscape
#' n_moves, number of timesteps
#' Requires:
#' tidyverse

move_host <- function(hid, hosts, coords, n_moves){
  ## grab host
  host_start <- filter(hosts, host_id %in% hid)
  
  ## grab landscape size for torus wrap
  max_x <- max(coords$x)
  max_y <- max(coords$y)
  
  ## random moves
  xmoves <- sample(c(-1, 0, 1), size = n_moves, replace = TRUE)
  ymoves <- sample(c(-1, 0, 1), size = n_moves, replace = TRUE)
  
  ## calculate x positions
  xpos <- cumsum(c(host_start$x, xmoves))
  
  ## calculate y positions
  ypos <- cumsum(c(host_start$y, ymoves))
  
  ## add positions to data
  host_moved <- tibble(host_id = hid,
                       xmove = c(NA, xmoves),
                       ymove = c(NA, ymoves),
                       pre_x = xpos,
                       pre_y = ypos,
                       time = 0:n_moves) %>%
    mutate(oob_x = (pre_x - 1) %/% max_x, ## calc out of bounds
           oob_y = (pre_y - 1) %/% max_y,
           x = pre_x - (max_x * oob_x), ## wrap around torus
           y = pre_y - (max_y * oob_y),
           id = str_c(x, y, sep = "_")) %>% ## generate ids
    left_join(select(coords, id, value), by = "id") %>% ## bring in parasite info
    mutate(parasite_gain = value * rbinom(n = n(), size = 1, prob = 0.5),
           parasite_burden = cumsum(parasite_gain)) %>% ## attach parasites
    filter(time == n_moves)
  
  ## return sim data
  return(host_moved)
}
```

Now testing it:

```{r}

test_sims <- replicate(10, 
                       parasite_sim(landscape = test_ls, n_hosts = 48, n_moves = 300),
                       simplify = FALSE)

test_sims_df <- bind_rows(test_sims, .id = "sim_id")
```

Works!

#### Generating Landscapes

Now lets make a bunch a varied landscapes:

```{r}

props <- rep(seq(1, 7, 2) / 10, 5)
clusters <- rep(seq(0, 20, 5), each = 4)

ls_list <- map2(
  .x = props,
  .y = clusters,
  ~fast_ls_perc(size = 48, prop = .x, cluster = .y),
  .progress = TRUE
)
```
