---
title: "sdp_parasite_sim_testing_p2_nb"
output: html_document
---

---
title: "sdp_parasite_sim_testing_p1_nb"
output: html_document
---

### Parasite Simulation Testing Part 2

First, source the functions:

```{r}
library(tidyverse)
library(som.nn)
library(ape)
library(gganimate)
library(gifski)
library(transformr)
source("parasite_sim_functions.R")
```

Now read the landscapes

```{r}

ls_list <- read_rds("sim_output/temp_ls_list_20230215_173748.rds")

```

Now run 10 sims per landscape

```{r}

test_sims <- map_df(ls_list, ~parasite_sim(.x, n_hosts = 96, n_moves = 300, n_reps = 10), .id = "sim_id", .progress = TRUE)
```

Now summarize and plot

```{r}
test_summary <- test_sims %>%
  filter(time == 300) %>%
  select(!host_id) %>%
  group_by(sim_id, rep_id) %>%
  summarise(mean_burden = mean(parasite_burden),
            var_burden = var(parasite_burden),
            disp_burden = var_burden / mean_burden) %>%
  left_join(select(test_sims, prop, moran, sim_id) %>% unique(), 
            by = "sim_id")

sim_plot <- ggplot(data = test_summary, aes(x = moran, y = disp_burden, color = factor(prop))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Moran's I (Clustering)", y = "Parasite Burden Dispersion") +
  scale_color_viridis_d(name = "Parasite \nProportion") +
  theme_bw() +
  theme(legend.position = "bottom")

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")

ggsave(filename = str_c("sim_output/temp_sim_plot_", 
                         timestamp,
                         ".png"), plot = sim_plot, width = 8, height = 6, units = "in")

```

#### Individual Variation

What if individual movement tendencies vary?

```{r}
test_sims_iv <- map_df(ls_list, 
                    ~parasite_sim(.x, n_hosts = 96, 
                                  n_moves = 300, n_reps = 10,
                                  movetypes = tibble(type = c(0.1, 0.3, 0.5, 0.7),
                                                     prop = rep(0.25, 4))),
                    .id = "sim_id", 
                    .progress = TRUE)


test_summary_iv <- test_sims_iv %>%
  filter(time == 300) %>%
  select(!host_id) %>%
  group_by(sim_id, rep_id) %>%
  summarise(mean_burden = mean(parasite_burden),
            var_burden = var(parasite_burden),
            disp_burden = var_burden / mean_burden) %>%
  left_join(select(test_sims, prop, moran, sim_id) %>% unique(), 
            by = "sim_id")

test_summaries <- test_summary %>%
  mutate(model = "No Individual Variation") %>%
  bind_rows(test_summary_iv %>%
              mutate(model = "Individual Variation")) %>%
  mutate(model = model %>%
           as.factor() %>%
           fct_relevel("No Individual Variation", "Individual Variation"))

iv_plot <- ggplot(data = test_summaries, aes(x = moran, y = disp_burden, color = factor(prop))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "Moran's I (Clustering)", y = "Parasite Burden Dispersion") +
  scale_color_viridis_d(name = "Parasite \nProportion") +
  facet_wrap(vars(model)) +
  theme_bw() +
  theme(legend.position = "bottom")

iv_plot

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")

ggsave(filename = str_c("sim_output/temp_iv_plot_", 
                         timestamp,
                         ".png"), plot = iv_plot, width = 10, height = 6, units = "in")

## summarize by individual
ind_summary <- test_sims_iv %>%
  filter(time == 300) %>%
  group_by(sim_id, rep_id, inactivity) %>%
  summarise(mean_burden = mean(parasite_burden),
            var_burden = var(parasite_burden),
            disp_burden = var_burden / mean_burden) %>%
  left_join(select(test_sims, prop, cluster, sim_id) %>% unique(), 
            by = "sim_id")

ind_plot <- ggplot(ind_summary, aes(x = inactivity, y = mean_burden, color = factor(prop))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(vars(cluster), nrow = 1) +
  scale_color_viridis_d(name = "Parasite \nProportion") +
  scale_x_continuous(name = "Probability of Not Moving", sec.axis = sec_axis(~ . , name = "Cluster Factor", breaks = NULL, labels = NULL)) +
  labs(y = "Mean Parasite Burden") +
  theme_bw() +
  theme(legend.position = "bottom")

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")

ggsave(filename = str_c("sim_output/temp_ind_plot_", 
                         timestamp,
                         ".png"), plot = ind_plot, width = 12, height = 6, units = "in")
  

```

#### Plotting Landscapes

Let's make a nice plot of the 20 landscapes

```{r}

ls_list_coords <- map_df(ls_list,
                         function(x){
                           bind_cols(x$coords, x$ls_stats)
                         },
                         .id = "ls_id")

## Second axis code from LizLaw
## https://stackoverflow.com/questions/11353287/how-do-you-add-a-general-label-to-facets-in-ggplot2

ls_plots <- ggplot(data = ls_list_coords) +
  geom_tile(aes(x = x, y = y, fill = factor(value), color = factor(value))) +
  scale_fill_manual(values = c("white", "salmon"), guide = "none") +
  scale_color_manual(values = c("white", "salmon"), guide = "none") +
  coord_fixed() +
  facet_grid(rows = vars(cluster), cols = vars(prop)) +
  scale_y_continuous(name = "", sec.axis = sec_axis(~ . , name = "Cluster Factor", breaks = NULL, labels = NULL)) +
  scale_x_continuous(name = "", sec.axis = sec_axis(~ . , name = "Parasite Proportion", breaks = NULL, labels = NULL)) +
  theme_bw() +
  theme(axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())

ls_plots

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")

ggsave(filename = str_c("sim_output/temp_ls_plots_", 
                         timestamp,
                         ".png"), plot = ls_plots)

moran_plot <- ggplot(data = ls_list_coords %>% 
                       select(!c(x, y, id, weight, value)) %>%
                       unique(),
                     aes(x = cluster, y = moran, color = factor(prop))) +
  geom_jitter(width = 0.5) +
  geom_smooth(method = "lm", se = FALSE, alpha = 0.2) +
  scale_color_viridis_d(name = "Parasite Proportion") +
  labs(x = "Cluster Factor", y = "Moran's I") +
  theme_bw()

moran_plot

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")
ggsave(filename = str_c("sim_output/temp_moran_plot_", 
                         timestamp,
                         ".png"), plot = moran_plot)

```

#### Plotting Simulations

Make some nifty animations!

```{r}

sample_ls <- ls_list[[10]]
sample_sim <- test_sims %>% filter(sim_id == 9, rep_id == 1)

ls_anim <- ggplot(data = sample_sim) +
  geom_tile(data = sample_ls$coords, aes(x = x, y = y, fill = factor(value))) +
  scale_fill_manual(values = c("white", "salmon"), guide = "none") +
  geom_point(aes(x = x, y = y), alpha = 0.5) +
  transition_time(time) +
  ease_aes('linear') +
  labs(title = "Timestep: {frame_time}") +
  theme_minimal() +
  theme_bw() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.grid = element_blank())

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")
anim_save(filename = str_c("sim_output/temp_ls_anim_", 
                         timestamp,
                         ".gif"), animation = ls_anim)


hist_anim <- ggplot(data = sample_sim) +
  geom_histogram(aes(x = parasite_burden), binwidth = 1, center = 0) +
  scale_y_continuous(limits = c(0, max(sample_sim$host_id))) +
  scale_x_continuous(limits = c(-1, max(sample_sim$parasite_burden))) +
  transition_time(time) +
  ease_aes('linear') +
  labs(title = "Timestep: {frame_time}", x = "Parasite Burden", y = "# Hosts") +
  theme_bw()

timestamp <- format(Sys.time(), format = "%Y%m%d_%H%M%S")
anim_save(filename = str_c("sim_output/temp_hist_anim_", 
                         timestamp,
                         ".gif"), animation = hist_anim)

```
